The stopping power of hydrogen in hafnium and \\ the importance of relativistic $4f$ electrons

### 2020-05-11 response.tex

%\reply{Thank you for the comment that helps us to clarify this point. Our SLPA calculations use the relativistic wave functions and binding energies for the bound electron contribution to the stopping power. The three curves in Fig. 3 and the curve in Fig. 4 are relativistic in this sense. We have added an explicit mention about the use of relativistic wave functions and binding energies in page 4 lines 200-201, and also this comment in page 4, lines 287-289:} \\ {\small ''Bound $1s$-$4f$ electrons (relativistic wave functions and binding energies) are calculated with the SLPA and shown separately in Fig. 3''}
\reply{We thank the referee for her/his comment since it will help us 
clarify this point. Our SLPA calculations use relativistic wave 
functions and binding energies for the bound electron contribution to 
the stopping power. In this sense, the three curves in figure 3 and the 
solid curve in Fig. 4 are called relativistic. We have added explicit 
mention of the use of relativistic wave functions and binding energies 
on page 4 lines 244-246 and lines 283-286.}

% ----------------------------------------------------------------------
\reviewer{3. If all the results of Fig. 3 were calculated for the 
relativistic description, it would be also interesting to show the
results obtained with the non-relativistic description in order to
demonstrate the importance of this correction.}

%\reply{The SLPA calculations for the bound electrons uses the Hartree-Fock wave funtions and binding energies, which are available in the tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54). The errors in the non-relativistic calculations worsen for higher Z atoms, and this is the reason why these tables do not extend any further than Z=54.  The fail of non-relativistic values of binding energies is shown in Fig. 2. We have added a comment on this in page 3, lines 214-228.} \\ {\small ``For more detail, Fig. 2 (b) shows  the relative errors with respect to the experimental  values. This figure shows clearly that the relativistic corrections are very important to describe the atomic structure of hafnium, even for the outer shells. It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''} \\ {\color{red} When dealing with stopping in targets with Z>54 we need to face our own calculation of the wave functions and binding energies. And we do it within the relativistic approximation to the best of our possibilities. However, one can wonder if such discrepancy in the non-relativistic results affects the stopping power. First, it will only affect the bound electron contribution to the total stopping. Second, at high energies (i.e. 1 MeV) the results converge to Bethe limit for any density of electrons as far as the correct number of target electrons is included. Having clarified this, we include in the figure \ref{NR} here, the total stopping using the non-relativistic values we calculate using the autostructure code. As can be noted, at sufficiently high energies it is the same, but differs around the stopping and at lower impact energies. We have included in this figure and in Fig. 4 of the manuscript the DPASS results by Sigmund.}
\reply{The SLPA calculations for the bound electrons uses the Hartree-
Fock wave functions and binding energies, which are available in the 
tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54). The 
errors in the non-relativistic calculations worsen for higher Z atoms, 
and this is the reason why these tables do not extend any further than 
Z=54. The fail of the non-relativistic description of the binding 
energies is shown in Fig. 2. We have added a comment on this on page 3, 
lines 220-230. When dealing with stopping in targets with Z>54, we need 
to compute the wave functions and binding energies ourselves. And we do 
it within the relativistic approximation to the best of our 
possibilities. However, one can wonder if such a discrepancy in the non-
relativistic results affects the stopping power. First, it will only 
affect the bound electron contribution to the total stopping. Second, at 
high energies (i.e., 1 MeV), the results converge to the Bethe limit for 
any density of electrons as far as the correct number of target 
electrons is included. Having clarified this, we added in figure 3 of 
this reply, the total stopping using the non-relativistic values we 
calculated using the AutoStructure code. As can be noted, at 
sufficiently high energy values, the results agree, but they start to 
differ around the stopping and at lower impact energies. We have
included in this figure, and in Fig. 4 of the manuscript, the DPASS
results by Sigmund.}

%-----------------------------------------------------------------------
\begin{figure}[!t]
\centering
\includegraphics[width=12.cm]{Fig05.eps}
\caption{Total stopping cross section. Present results using
relativistic and non-relativistic solutions for hafnium atom. 
Curves and symbols explained inside the figure.}
\label{NR}
\end{figure}
% ----------------------------------------------------------------------
\reviewer{
4. Furthermore, it would also be meaningful to show the results from
the ML(FEG) calculation alone in order to identify better the
contribution of the bound electrons to the SP.}

%\reply{Thank you for this suggestion. We have added the curves for FEG and bound electrons to Fig. 3, and the corresponding comments in the text and in the figure caption.}
\reply{Thank you for this suggestion. We have added the curves for FEG and bound electrons to Fig. 3 of the manuscript, and the corresponding comments in the text and figure caption.}

% ----------------------------------------------------------------------
\reviewer{
5. When comparing the new and existing experimental data, clarify what
experimental technique was used in Ref. 7 to measure the stopping
power.}

\reply{Good suggestion, thank you. Sirotinin et al. used the 
backscattering method for these measurements. We have made reference to 
the method implemented in the Introduction (line 70) and Section IV 
(line 305).}

% ----------------------------------------------------------------------
\reviewer{
6. The data by Sirotonin et al. plotted in Fig. 4 seem to suggest that
the stopping power maximum would be around 100-200 keV, while the
maximum predicted in the present work by the theoretical calculations
is located at slightly smaller energies. Please comment on this point
and discuss possible reasons for this discrepancy: What are the main
limitations in the theoretical model that can be causing this energy
shift?}

\reply{The stopping maximum is very sensitive to the theoretical model and to the experimental method. Sirotinin paper mention "possible reasons for the disagreement of the measurement results near the E(E) curve maximum". 
%The original paper by Sirotinin says: "The possible reasons for the disagreement of the measurement results near the E(E) curve maximum are mainly related to the inaccuracy in determining thickness, to nonuniformity of the thickness, and the neglect of the effect of the chemical composition and structure of the thin films used in the measurements"
For targets with many different sets of data (i.e. H in Ag, Al, Au, Fe, Cu, Si, and others) it is usual to note great dispersion of values around the maximum. In the case of Hf, only one set of data is available and it is perfectly matched by SRIM semi-empirical curve. SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data. One of the goals of this work is to call the attention on possible differences between theory and the existing values, and to remark the necessity of more experimental data in certain energy regions, mainly, the stopping maximum and below. We have changed the comments in page 6 lines 328-336 and lines 364-366 as follows:} \\ {\small ''Future experiments  would be important to complete this study on the stopping of Hf for protons. In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available. SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data. Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e. $50-200$ keV) and at low impact energies''} \\ {\small ''Future experiments for impact energies around the stopping maximum and in the low energy region would be important to complete this study.''} \\ {\color{red} In relation with the referee comment about the maximum of the stopping, in Fig. 4 we have added the comparison with other two theoretical results: CASP and PASS. The following comment has been included in page 6, lines 315-322:} \\ {\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig. [4]. However, the impact energy for the maximum seem to agree between our curve and DPASS, although DPASS is $10 \%$ below ours. Instead, CasP predicts a stopping maximum $10 \%$ above our value and a completely different shape at lower energies.''}

% ----------------------------------------------------------------------
\reviewer{
7. When reviewing on the experimental work performed on Hf oxides in
the introduction, note that there is a quite recent work by Roth and
co-workers published in Phys. Rev. Lett. 119, 163401 (2017), in which
the SP of protons is precisely measured in this target.}

\reply{We thank the referee for suggesting this missing reference. We have included it in the revised manuscript.}

% ----------------------------------------------------------------------
\reviewer{Finally, there are few misprints in the generated pdf file:}

\reviewer{- Y-axis labels in Figs. 3 an 4 are unreadable }

\reply{We have changed to larger prints  and have written explicitly "stopping cross section" -Y- and "Impact energy" -X- instead of S(E) and E. }

\reviewer{- Page 3,2nd column, 2nd full paragraph: "Too assess"}

\reply{Corrected.}

\reviewer{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
"unfilled circles"}

\reply{Corrected.}

\reply{We have also made some improvements in the English redaction of the manuscript.}
